---
title: "Day 9 Homework"
output: html_document
---

1. Write a function to do a bunch of clean-up and formatting to your data.

```{r}
library(reshape2)
library(dplyr)

mySP <- read.csv("SPTimePeriods.csv")
mySP$RelID <- 1:nrow(mySP)
mySP$RelID <- paste("Rel", 1:nrow(mySP), sep = "")

melt.data <- function(data) {
  mySP.melt <- melt(data, id.vars = c("RelID", "ID", "Relationship"), value.name = "relationship.type", variable.name = "time.period")
}


```


2. Make sure this function works on your data.

```{r}
melt.data(mySP)
mySP.melt

```


3. Get some subset of your data, or find a similar-but-not-identical data set.  Change up the column order.  Re-run your function on this new data.  Does it work?

```{r}
#subsetting data - changing the column order doesn't affect the data since the columns are all similarly constructed


mySP.subset <- select(mySP, ID:T5)
mySP.subset$RelID <- 1:nrow(mySP.subset)
mySP.subset$RelID <- paste("Rel", 1:nrow(mySP), sep = "")
head(mySP.subset)

# running the function on the new dataset:
melt.data2 <- function(data) {
  mySP.melt2 <- melt(data, id.vars = c("RelID", "ID", "Relationship"), value.name = "relationship.type", variable.name = "time.period")
}

melt.data2(mySP.subset)
mySP.melt2

#it seems to work
```


4. Fix your function to be able to work on the new data.

5. Write a function to do some kind of summary stat, and use the `dplyr` functions to run this function over groups of your data.

```{r}

max.reltype <- function(data) {
  group_by(data, time.period)
  output <- summarize(data, maxrel = max(relationship.type))
}

max.reltype(mySP.melt)
head(mySP.melt)

output

#the only way I can do a summary statistic that is meaningful is to recast the data into the distributions - I tried to get the maximum value of the relationship.type column of the .melt data as the output as a test, but it continues to give me errors saying that the .data value is not in the summarize() part of the code. Not sure how to fix it.

```
