---
title: "Day 5 Homework"
output: html_document
---

0. Start out by getting your data into "long" format.  This will help with some of the following.
```{r}
install.packages("ggplot2")
library(reshape2, "ggplot2")
#read data into R
setwd("~/R/Homework")
mySP <- read.csv("SPTimePeriods.csv")
#create a relationship ID variable to help make sure that data is not aggregated across individuals when melted and aggregated differently
mySP$RelID <- 1:nrow(mySP)
mySP$RelID <- paste("Rel", 1:nrow(mySP), sep = "")
head(mySP)
#melt data to form a "long" version
mySP.melt <- melt(mySP, id.vars = c("RelID", "ID", "Relationship"), value.name = "relationship.type", variable.name = "time.period")
summary(mySP.melt)
head(mySP.melt)
```

1. Pick a single numerical variable from your data.  Use `ggplot` to plot a histogram and a density plot (separately).
   - Try some different values for `binwidth` for the histogram.
   - Try some different values for `adjust` for the density plot.

```{r}
mybaseplot <- ggplot(mySP.melt, aes(x = relationship.type))
mybaseplot + geom_histogram(binwidth = 1)
mybaseplot + geom_histogram(binwidth = .5)


mybaseplot + geom_density(binwidth = 1)
mybaseplot + geom_density(binwidth = .2)
```

2. How does the distribution look?  Try transforming the variable and try to get it to look more normal (i.e., more Gaussian).  If it already looks normal, do something to it to make it look non-normal.

#the relationship looks skewed to the right, taking the log of the relationship type makes it marginally more normally distributed.

```{r}
mybaseplot <- ggplot(mySP.melt, aes(log(relationship.type)))
mybaseplot + geom_histogram(binwidth = .2)

#another example:
#minpair <- read.csv("scottdata/CognitionPaperFinalData.csv")
#head(minpair)
#library(ggplot2)
#ggplot(minpair, aes(log(WithinMinPairs))) + geom_histogram()  + geom_vline(xintercept = 1000, color = "red")
#ggplot(minpair, aes(log(WithinMinPairs))) + geom_histogram(aes(fill = Corpus)) + scale_fill_brewer(palette = "Set1") - allows you to choose the color palette for the plot via color brewer (http://colorbrewer2.org/)

colors()
```

3. Is there a factor that you can use to break up this variable into groups of observations?  If not, reshape your data so that there is, or create a factor if necessary.  Then, do the following:
   - Plot it as a density plot, mapping `color` to the factor
   - Same, but try using `fill` instead of `color`
   - Same, using `linetype` instead of `color`
   - Plot the data as a boxplot, with the factor as `x` and the numerical variable as `y`

```{r}
mybaseplot <- ggplot(mySP.melt, aes(x = time.period))
mybaseplot + geom_density(aes(color = time.period))
mybaseplot + geom_density(aes(fill = time.period))
mybaseplot + geom_density(aes(linetype = time.period))
ggplot(mySP.melt) + geom_boxplot(aes(x = time.period, y = relationship.type))

```


4. Now let's look at breaking up a histogram across a factor.  With the same variables you used in #3:
   - Plot a histogram, but use faceting to display them, one above the other.
   - Try mapping the factor to `color` or `fill`, to see what that looks like.
   - Which one seems most useful or informative for your data, and why?

```{r}

mybaseplot + geom_histogram(aes(color = time.period)) + facet_grid(relationship.type ~ .)
mybaseplot + geom_histogram(aes(fill = time.period)) + facet_grid(relationship.type ~ .)
mybaseplot + geom_histogram(aes(color = time.period))
mybaseplot + geom_histogram(aes(fill = time.period))

#facet_wrap takes a long row of pltos and tries to wrap them to make as square a visualization as possible
#facet_grid - you have two dimensions and tell it how you want each of the dimensions to be - best for two variables.

```

#the facet grid with the fill provides the most information becuase it allows you to easily see the distribution of relationship types over the time periods

5. Now find two numerical variables.  You may need to reshape your data.  For example, if you have one numerical variable and a "condition" factor, you could use `dcast` to spread that variable across multiple columns (one per level of the factor).  Once you have two numerical variables:
   - Create a scatterplot, with one variable mapped to `x` and the other to `y`.
   - Plot a "loess" smoother on top of the scatterplot
   - Plot a regression line on top of the scatterplot (instead of the loess)
  
```{r}
?dcast
head(mySP.melt)
mySP.recast <- dcast(mySP.melt, RelID + ID + Relationship ~ time.period, value.var = "relationship.type")
head(mySP.recast)
scatter <- ggplot(mySP.recast, aes(x = T6, y = T7)) 
scatter + geom_point()
scatter + geom_point() + geom_smooth(method = "loess")
scatter + geom_point() + geom_smooth(method = "lm")

```


6. Are there any other factors that split up the variable plotted in #5?  If so try the following:
   - Map the factor to `color`
   - Facet by the factor

```{r}
scatter + geom_point(aes(color = RelID))
scatter + geom_point() + facet_grid(T9 ~ .)

```

7. Output all of the plots above as a single PDF.

```{r}
pdf("myplot.pdf", width = 10, height = 4)
Q1.Hist <- ggplot(mySP.melt, aes(x = relationship.type)) + geom_histogram(binwidth = 1)
Q1.Dens <- ggplot(mySP.melt, aes(x = relationship.type)) + geom_density(binwidth = .2)
Q2.hist <- ggplot(mySP.melt, aes(log(relationship.type))) + geom_histogram(binwidth = .2)
Q3.dencol <- ggplot(mySP.melt, aes(x = time.period)) + geom_density(aes(color = time.period))
Q3.denfill <- ggplot(mySP.melt, aes(x = time.period)) + geom_density(aes(fill = time.period))
Q3.denline <- ggplot(mySP.melt, aes(x = time.period)) + geom_density(aes(linetype = time.period))
Q3.box <- ggplot(mySP.melt) + geom_boxplot(aes(x = time.period, y = relationship.type))
Q4.facetcol <- ggplot(mySP.melt, aes(x = time.period)) + geom_histogram(aes(color = time.period)) + facet_grid(relationship.type ~ .)
Q4.facetfill <- ggplot(mySP.melt, aes(x = time.period)) + geom_histogram(aes(fill = time.period)) + facet_grid(relationship.type ~ .)
Q4.histcol <- ggplot(mySP.melt, aes(x = time.period)) + geom_histogram(aes(color = time.period))
Q4.histfill <- ggplot(mySP.melt, aes(x = time.period)) + geom_histogram(aes(fill = time.period))
Q5.scatter <- ggplot(mySP.recast, aes(x = T6, y = T7)) + geom_point()
Q5.scatterloess <- ggplot(mySP.recast, aes(x = T6, y = T7)) + geom_point() + geom_smooth(method = "loess")
Q5.scatterlm <- ggplot(mySP.recast, aes(x = T6, y = T7)) + geom_point() + geom_smooth(method = "lm")
Q6.scattercol <- ggplot(mySP.recast, aes(x = T6, y = T7)) + geom_point(aes(color = RelID))
Q6.scatterfacet <- ggplot(mySP.recast, aes(x = T6, y = T7)) + geom_point() + facet_grid(T9 ~ .)
dev.off()

```

8. Is there a better type of visualization for your data?  What numbers would you like to be able to visualize better?  Be as explicit as possible, and write some "pseudo-code" to ilustrate what you think you would like.

#a network diagram would be better for my data. Specifically, I would like to be able to visualize the relationships between the individuals. So each node would represent a person and the line between them (or edge) would represent the relationship. edges would be colored based on the type of relationship and there would be a visualization for each time period. To do so, I would need to use igraph, the network analysis package.

#some potential sample code could be the following. The graph.adjacency command seems to display the network from an adjacency matrix format. I would neet to confirm that the edgearray format my data is in qualifies as an adjacency matrix as defined by R.
```{r}
#install.packages(igraph)
#library(igraph)
#graph.adjacency(mySP, mode ="directed", weighted = NULL, diag = TRUE, add.colnames = NULL, add.rownames = NA)

#alpha allows for transparency in the color but it is not an aesthetic, so it'a parameter of density usually. e.g. geom_density(aes(fill = variable), alpha = .3)
#themes can be changed to alter the appearance of the image/plot

```

9. Save your Markdown document that answers all of these questions (with code) as well as your PDF, and commit them as changes to your git homework repository.  Push those changes to GitHub by Monday morning at 9am.


