---
title: "Day 10 Homework"
output: html_document
---

1. Find some relationship in your data (e.g., correlation, comparison of means, etc.) that is "significant" (don't worry too much about what that might mean).

```{r}

library(reshape2)
library(dplyr)

# get melted dataset

mySP <- read.csv("SPTimePeriods.csv")
mySP$RelID <- 1:nrow(mySP)
mySP$RelID <- paste("Rel", 1:nrow(mySP), sep = "")


mytest <- t.test(mySP$T3, mySP$T4, paired = TRUE)
pval <- mytest$p.value

```


2. Get an estimate of the effect size (e.g., the difference between means, or the correlation).

```{r}
nsim <- 1000

results.table2 <- data.frame(diff = rep(NA, nsim), pval = NA)

for(i in 1:nsim) {
  simtest <- t.test(mySP$T3, mySP$T4, paired = TRUE)
  simdiff <- simtest$estimate
  pval <- simtest$p.value
  results.table2[i, ] <- c(simdiff, pval)
}

results.table2

```


3. Simulate data with this effect size (either using mvrnorm or adding a simulated effect), but with a much smaller data sample.  In other words, simulate an underpowered study.

```{r}

#see above
```


4. Examine the amount of effect-size inflation via the "significance filter" by looking at the effect sizes from the simulations, for just the sims where the effect turned out significant.

```{r}

##don't know

alpha <- .05
results.sig2 <- filter(results.table2, pval < alpha)
library(ggplot2)

ggplot(results.table2, aes(diff)) + geom_histogram(aes(fill = significant)) + scale_fill_brewer(palette = "Set1") + geom_vline(xintercept = .3, linetype = 2) + facet_grid(significant ~ .)

```

