---
title: "Day 9 Homework"
output: html_document
---

1. Write a function to do a bunch of clean-up and formatting to your data.

```{r}
library(reshape2)
library(dplyr)

mySP <- read.csv("SPTimePeriods.csv")

melt.data <- function(data) {
  data$RelID <- paste("Rel", 1:nrow(data), sep = "")
  mySP.melt <- melt(data, id.vars = c("RelID", "ID", "Relationship"), value.name = "relationship.type", variable.name = "time.period")
  mySP.melt
}


```


2. Make sure this function works on your data.

```{r}
melt.data(mySP)
mySP.melt

```


3. Get some subset of your data, or find a similar-but-not-identical data set.  Change up the column order.  Re-run your function on this new data.  Does it work?

```{r}
#subsetting data - changing the column order doesn't affect the data since the columns are all similarly constructed


mySP.subset <- select(mySP, ID:T5)
head(mySP.subset)

# running the function on the new dataset:
melt.data2 <- function(data) {
  data$RelID <- paste("Rel", 1:nrow(data), sep = "")
  mySP.melt2 <- melt(data, id.vars = c("RelID", "ID", "Relationship"), value.name = "relationship.type", variable.name = "time.period")
  mySP.melt2
}

melt.data2(mySP.subset)
mySP.melt2

#it seems to work
```


4. Fix your function to be able to work on the new data.

5. Write a function to do some kind of summary stat, and use the `dplyr` functions to run this function over groups of your data.

```{r}

# NOT with a Function

## get the mean for each time period
step1 <- mySP.melt %>%
  group_by(time.period) %>%
  summarize(mean = mean(relationship.type))
step1

## merge into the melted dataframe
### mySP.melt <- merge(mySP.melt, step1)
### head(mySP.melt)

## include it in the casted version
mySP.cast2 <- dcast(mySP.melt, time.period ~ relationship.type, fun.aggregate = length)

mySP.cast2 <- merge(mySP.cast2, step1)
mySP.cast2

# WITH a Function

mean <- function(data) {
  group_by(data, time.period)
  step1 <- summarize(data, mean = mean(relationship.type))
  step1
  result <- dcast(data, time.period ~ relationship.type)
  result <- merge(result, step1)
  result
}

mean(mySP.melt)

```


#Notes from discussion:

```{r}
get("x") #lets you recover/recall the object with the name of "x" - works to recall a variable with a string name
```


