---
title: "Day 8 Homework"
output: html_document
---

1. Try to recreate the results of Homework 6 (or your intended results) using the `dplyr` functions.

```{r}
library(reshape2)
library(dplyr)
mySP <- read.csv("SPTimePeriods.csv")
mySP$RelID <- 1:nrow(mySP)
mySP$RelID <- paste("Rel", 1:nrow(mySP), sep = "")
head(mySP)
mySP %>%
  melt(mySP, id.vars = c("RelID", "ID", "Relationship"), value.name = "relationship.type", variable.name = "time.period") %>%
  group_by(time.period) %>%
  mySP.cast <- dcast(mySP.melt, time.period ~ relationship.type)

head(mySP.cast)
```


2. Find another kind of analysis/transformation/summary stat that you want to apply across multiple groups in your data. Use the functions from `reshape2` and `dplyr` to get these results.

```{r}

summarize(mySP.melt %>% group_by(time.period), mean = mean(relationship.type))

```


3. For one more practice, try another example of something like #2 (find another stat or grouping or something).

```{r}

summarize(mySP.melt %>% group_by(time.period), max = max(relationship.type), min = min(relationship.type), mean = mean(relationship.type))

```

